Initial Exploration

### 2019-02-22 build.Rmd
update

#Overview
This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.

Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html

Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html

>>>>

#Overview
This page links to sub-pages that describe how various components used in the project were built. Follow the links for details.

Extracting TAs' comments: https://adanieljohnson.github.io/default_website/docxcommentextraction.html

Structure of development dataset; data import & validation: https://adanieljohnson.github.io/default_website/dataimport.html

Current Naive Bayes classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html

### 2019-02-22 dataimport.Rmd
update

#Overview
This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.

Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html

Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html

@@

#Describe an Element

#Describe an Element

Preliminary exploration of the dataset is described here: https://adanieljohnson.github.io/default_website/initial_exploration.html

Rationale and initial steps in development of the Classifier are described here: https://adanieljohnson.github.io/default_website/bayes_history.html

Code for the current version of the Naive Bayes Classifier is located here:
https://adanieljohnson.github.io/default_website/NB_2r2c.html

@@

@@

##Initial Library and Data Import
```{r Libraries_1}
# Call in tidyverse, other required libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

@@

#Part 3. Post-Import Checks
Given some data points may be entered incorrectly by instructors or students, these post-import data checks help ensure data have been properly coded and columns contain valid entries only. (This process can be automated using the ```janitor``` data cleaning package.)

```{r}
#"Course" column should only have entries that correspond to 113, 114, 214, NA. 
course<-unique(base_data$course)

#Entries in "TAs" column should match the anonymous IDs that correspond to the TAs assigned to the relevant courses in the semester being analyzed. 
#If there are incorrect or extra anonymous IDs in this list, the mostly likely source of the error is improper de-identification of the Excel file prior to import.
ta<-unique(base_data$ta)

#"Lab" should only have entries that match the allowed limited keywords in codebook.
lab<-unique(base_data$lab)

#It is impractical to validate all randomized identifiers. Instead look at the NUMBER of unique anonymous student IDs. Quantity should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
student<-unique(base_data$student)

#Check that the list of "code.subject" topics here matches allowed terms in codebook.
subject<-unique(base_data$code.subject)

#Check that the list of "code.structure" terms extracted from the dataset and displayed here matches allowed terms in codebook.
structure<-unique(base_data$code.structure)
```
####\

>>>>

The dataset being used to develop this project is an Excel table containing 10,808 TA comments extracted from student lab reports submitted during Spring 2018. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

@@

Data were anonymized in Excel, replacing names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.

@@

#Pre-Checks Before Import
Prior to import, review the data table **in Excel** and check that:

1.  Student and TA names have been recoded using pre-assigned anonymous IDs.
2.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd" [LINK]
3.  Data table headers match those listed below. The same list is in the file "code_column_names.rmd" [LINK]

@@

#Initial Data Import
```{r Libraries_1}
# Call in tidyverse, other required libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

@@

#Post-Import Checks
Given some data points may be entered incorrectly by instructors or students, these post-import data checks ensure data are properly coded and columns contain valid entries only. (This process can be automated using the ```janitor``` data cleaning package.)

```{r}
#"Course" column should only have entries that correspond to 113, 114, 214, NA. 
course<-unique(base_data$course)
```

```{r}
#Entries in "TAs" column should match the anonymous IDs that correspond to the TAs assigned to the relevant courses in the semester being analyzed. 
#If there are incorrect or extra anonymous IDs in this list, the mostly likely source of the error is improper de-identification of the Excel file prior to import.
ta<-unique(base_data$ta)
```

```{r}
#"Lab" should only have entries that match the allowed limited keywords in codebook.
lab<-unique(base_data$lab)
```

```{r}
#It is impractical to validate all randomized identifiers. Instead look at the NUMBER of unique anonymous student IDs. Quantity should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
student<-unique(base_data$student)
```

```{r}
#Check that the list of "code.subject" topics here matches allowed terms in codebook.
subject<-unique(base_data$code.subject)
```

```{r}
#Check that the list of "code.structure" terms extracted from the dataset and displayed here matches allowed terms in codebook.
structure<-unique(base_data$code.structure)
```
####\

### 2019-02-22 hiding/datastructure.Rmd
update

---
title: "Data Structure"
author: "Dan Johnson"
date: "2/22/2019"
output: html_document
---

@@

#Overview
This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.

Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html

Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html

#Describe an Element

#Describe an Element

#Overview
This page describes how the target dataset is structured.

How data are imported then validated for analysis is described here: https://adanieljohnson.github.io/default_website/dataimport.html

Rationale and initial steps in development of the Classifier are described here: https://adanieljohnson.github.io/default_website/bayes_history.html

Code for the current version of the Naive Bayes Classifier is located here:
https://adanieljohnson.github.io/default_website/NB_2r2c.html

#Target Dataset
The dataset being used to develop this project is an Excel table containing 10,808 TA comments extracted from student lab reports submitted during Spring 2018. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

>>>>

---
title: "DROPPED"
author: "Dan Johnson"
date: "2/22/2019"
output: html_document
---

@@

#Overview
The dataset being used to develop this project is an Excel table containing 10,808 TA comments extracted from student lab reports submitted during Spring 2018. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

### 2019-02-22 initial_exploration.Rmd
update

---
title: "Initial Explorations"
output: html_document
---

@@

The first question to answer before making deeper comparisons is **are TAs' comments obviously different between sub-categories?** To that end, I ran three different analyses of word usage in TA comments.

*  Pearson correlations for word use frequencies. This compares words found in all Subject sub-categories
*  Word frequency differences based on log-odds ratios. This highlights terms that are the most different between sub-categories.
*  Unique n-grams mapping. This identifies text strings, phrases, and rare terms that are unique (or nearly unique) to one sub-category.

@@

#Pearson Correlations of Word Frequencies
##Are Shared Words Unequally Distributed?
This analysis takes all comments from base_data, tokenizes and filters stopwords, then counts and sorts fractional frequencies for each word in each category. **NOTE**: this analysis only compares word use for terms present in ALL sub-categories. If a word is MISSING from **any subcategory**, the data row for that word is pruned from the dataset. This eliminates all unique words in favor of words found in all sub-categories.

@@

#Log-Odds Ratios
##Which Words Are Shared LEAST Between Categories?
Unlike the preceding analysis, log-odds ratio comparison does not eliminate unshared terms. Instead, it identifies which words are MOST likely to be used in one sub-category, but not another. For simplicity I reduced the analysis to just the four most common comment subjects. This has been the most informative preliminary visualization I have found so far.

@@

#N-Grams Mapping 
##Which Terms, Phrases Are Unique to a Sub-Category?

```{}
#Pseudo-code
Import comments
Tokenize the comments as 1-, 2, 3-grams
Add counts within subset, calculate frequency
Split by subcategory
Align 4 subcategories, highest to lowest

>>>>

---
title: "Initial Exploration"
output: html_document
---

@@

#Overview
The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** To that end, I ran three different analyses of word usage frequency.

*  __Pearson correlations for word use frequencies__. This only compares words found in all of the Subject sub-categories
*  __Word frequency differences based on log-odds ratios__. This highlights terms that are the most different between sub-categories.
*  __N-grams mapping__. This identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category.

@@

#Pearson Correlations of Word Frequencies
##Are Shared Words Used At Different Frequencies?
This analysis takes all comments from base_data, tokenizes and filters stopwords, then counts and sorts fractional frequencies for each word in each category. **NOTE**: this analysis only compares word use for terms present in ALL sub-categories. If a word is MISSING from **any subcategory**, the data row for that word is pruned from the dataset. This eliminates all unique words in favor of words found in all sub-categories.

@@

#Log-Odds Ratios of Word Frequencies
##Which Words Are Shared LEAST Between Categories?
Unlike the preceding analysis, log-odds ratio comparison does not eliminate unshared terms. Instead, it identifies which words are MOST likely to be used in one sub-category, but not another. For simplicity I reduced the analysis to just the four most common comment subjects. This has been the most informative preliminary visualization I have found so far.

@@

#N-Grams Mapping 
##Are Certain Terms and Phrases Unique to One Sub-Category?
One of our earliest observations was that the n-gram "basic criteria" only appeared in TA comments from three different Subject sub-categories. This suggests certain terms and n-grams might be eliminated

However it w  the relative frequency was much higher
```{}
#Pseudo-code
Import comments
Tokenize the comments as 1-, 2, 3-grams
Add counts within subset, calculate frequency
Split by subcategory
Align 4 subcategories, highest to lowest
